---
title: "R Notebook: SOC 5800 Team Project"
author: "Malcolm S. Townes, Jacob Eikenberry, Daniel Ferris, Nicholas Sokolis"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
    word_document: default
    html_notebook: default
    html_document: default
    github_document: default
    always_allow_html: yes
---

## Introduction
This is an R Notebook for a study that examines how formerly incarcerated individuals who successfully reintegrated cope with barriers to re-entry.  This study was conducted as a team project for the SOC 5800 Survey Design and Sampling course during the Fall 2019 semester at Saint Louis University.

## Project Set Up
The following code chunk enables the R Notebook to integrate seamlessly with the project organization format. This is normally included in the R Notebook to simplify file calls and enable file portability but it has been causing an error. To work around this problem, I've embedded the `here()` function where I enter a file path when necessary.

```{r setup-project, message = FALSE, warning = FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

## Load Dependencies
The following code chunk loads package dependencies required to perform the necessary tasks. Basic tasks include importing, reading, wrangling, and cleaning data; selecting a subset of the data; checking for unique observations; analyzing missing data; and performing various types of regression analyses. 
```{r load-dependencies, message = FALSE, warning = FALSE}
library(tidyverse) # loads the basic R packages
library(here) # enables file portability
library(readr) # functions for reading data
library(dplyr) # functions for data wrangling
library(janitor) # functions for data cleaning
library(naniar) # functions for analyzing missing data
library(expss) # functions for calculating on values
library(ggplot2) # functions for data visualizations
library(boot) # functions for regression analysis
library(generalhoslem) # Hosmer-Lemeshow test for binary and multinomial logistic models
library(ordinal) # functions for regression models for ordinal data
library(MASS) # functions for ordered logistic or probit regression
library(broom) # functions for tidying ordinal logistic regression models
library(gvlma) # functions for global validation of linear model assumptions
library(lmtest) # functions for testing linear regression models
library(ltm) # functions for latent trait models under Item Response Theory
library(leaps) # functions for regression subset selection
library(car) # companion to applied regression
library(aod) # functions to analyze overdispersed data counts and proportions
library(pscl) # contains function for pseudo R2 measures for logistic regression
library(ResourceSelection) # contains function for Hosmer-Lemeshow goodness of fit test
library(psy) # functions for various procedures used in psychometry
library(psych) # Procedures for Psychological, Psychometric, and Personality Research
```

## Load Raw Data
The following code chunk imports the raw data from the `csv` file.
```{r import-data, message = FALSE, warning = FALSE}
dataRaw <- read.csv(here("Data","DataRaw","SOC5800_Data_NumericValues_Raw_CSV.csv"), 
                      sep = ",", header = TRUE, fill = TRUE, dec = ".")
```

## Rename Variables
The following code chunk performs several actions to clean the raw data. It first renames the variables with descriptive camel case names. It then removes unused variables.
```{r clean-data, message = FALSE, warning = FALSE}
dataRaw %>%
  dplyr::rename(startDate = StartDate,
         endDate = EndDate,
         status = Status,
         IPaddress = IPAddress,
         progress = Progress,
         sessionDuration = Duration..in.seconds.,
         surveyCompleted = Finished,
         surveyDate = RecordedDate,
         responseID = ResponseId,
         locationLat = LocationLatitude,
         locationLong = LocationLongitude,
         consent = Q43,
         browserName = Q10_Browser,
         browserVersion = Q10_Version,
         opSyst = Q10_Operating.System,
         screenRes = Q10_Resolution,
         currentlyIncarcerated = Q1,
         currentlyDetained = Q2,
         residentialTreatment = Q3,
         involCommitment = Q4,
         gender = Q5,
         genderSelfDescribe = Q5_3_TEXT,
         transgender = Q7,
         ethnicityRace = Q6,
         ethnicityRaceOther = Q6_10_TEXT,
         typeHometown = Q8,
         religiousAffiliation = Q9,
         ReligiousOther = Q9_12_TEXT,
         ageNow = Q11,
         ageRelease = Q13,
         incarcerationYears = Q14.1_1,
         incarcerationMonths = Q14.2_1,
         educationLevels = Q15,
         whenHighestEd = Q16,
         relationshipStatus = Q17,
         householdSize = Q18,
         financialSupportInitial = Q19,
         financialSupportGovt = Q19_4_TEXT,
         financialSupportNonGovt = Q19_5_TEXT,
         finanicalSupportOther = Q19_6_TEXT,
         selfEmployment = Q20,
         incomeInitial = Q21,
         incomeLastYr = Q22,
         managingFinancially = Q23,
         savings = Q24,
         pssQ1 = Q41_1,
         pssQ2 = Q41_2,
         pssQ3 = Q41_3,
         pssQ4 = Q41_4,
         pssQ5 = Q41_5,
         pssQ6 = Q41_6,
         pssQ7 = Q41_7,
         pssQ8 = Q41_8,
         pssQ9 = Q41_9,
         pssQ10 = Q41_10,
         pssQ11 = Q41_11,
         pssQ12 = Q41_12,
         issFamilyQ1 = Q30_1,
         issFamilyQ2 = Q30_2,
         issFamilyQ3 = Q30_3,
         issFamilyQ4 = Q30_4,
         issFamilyQ5 = Q30_5,
         issFriendQ1 = Q31_1,
         issFriendQ2 = Q31_2,
         issFriendQ3 = Q31_3,
         issFriendQ4 = Q31_4,
         issFriendQ5 = Q31_5,
         programsUsed = Q32,
         importancePublicTrans = Q33_1,
         importanceHousing = Q33_2,
         importanceSNAP = Q33_3,
         importanceWIC = Q33_4,
         importanceTANF = Q33_5,
         importanceEITC = Q33_6,
         importanceJobTraining = Q33_7,
         importanceMedicare = Q33_8,
         importanceEducAid = Q33_9,
         aceQ1 = Q34_1,
         aceQ2 = Q34_2,
         aceQ3 = Q34_3,
         aceQ4 = Q34_4,
         aceQ5 = Q34_5,
         aceQ6 = Q34_6,
         aceQ7 = Q34_7,
         aceQ8 = Q34_8,
         aceQ9 = Q34_9,
         aceQ10 = Q34_10,
         helpDuring = Q35,
         helpDuringType = Q33,
         helpDuringOther = Q33_8_TEXT,
         helpAfter = Q36,
         helpAfterType = Q44,
         helpAfterOther = Q44_8_TEXT,
         gritQ1 = Q35_1,
         gritQ2 = Q35_2,
         gritQ3 = Q35_3,
         gritQ4 = Q35_4,
         gritQ5 = Q35_5,
         gritQ6 = Q35_6,
         gritQ7 = Q35_7,
         gritQ8 = Q35_8,
         stayInitial = Q36.1,
         stayInitialOther = Q36_6_TEXT,
         foodSecurityQ1 = Q37_1,
         foodSecurityQ2 = Q37_2,
         mostHelpful = Q38,
         greatestObstacle = Q39,
         helpKind = Q40,
         justiceInteraction = Q41
         ) -> dataRenamed

dataRenamed %>%
  dplyr::select (-c(status, progress, RecipientLastName, RecipientFirstName, RecipientEmail, ExternalReference)) -> dataAugmented
```

## Change Data Type
The following code chunk changes the data type for certain variables.
```{r change-data-types, message = FALSE, warning = FALSE}
dataAugmented$pssQ1 <- as.numeric(dataAugmented$pssQ1)
dataAugmented$pssQ2 <- as.numeric(dataAugmented$pssQ2)
dataAugmented$pssQ3 <- as.numeric(dataAugmented$pssQ3)
dataAugmented$pssQ4 <- as.numeric(dataAugmented$pssQ4)
dataAugmented$pssQ5 <- as.numeric(dataAugmented$pssQ5)
dataAugmented$pssQ6 <- as.numeric(dataAugmented$pssQ6)
dataAugmented$pssQ7 <- as.numeric(dataAugmented$pssQ7)
dataAugmented$pssQ8 <- as.numeric(dataAugmented$pssQ8)
dataAugmented$pssQ9 <- as.numeric(dataAugmented$pssQ9)
dataAugmented$pssQ10 <- as.numeric(dataAugmented$pssQ10)
dataAugmented$pssQ11 <- as.numeric(dataAugmented$pssQ11)
dataAugmented$pssQ12 <- as.numeric(dataAugmented$pssQ12)
dataAugmented$issFamilyQ1 <- as.numeric(dataAugmented$issFamilyQ1)
dataAugmented$issFamilyQ2 <- as.numeric(dataAugmented$issFamilyQ2)
dataAugmented$issFamilyQ3 <- as.numeric(dataAugmented$issFamilyQ3)
dataAugmented$issFamilyQ4 <- as.numeric(dataAugmented$issFamilyQ4)
dataAugmented$issFamilyQ5 <- as.numeric(dataAugmented$issFamilyQ5)
dataAugmented$issFriendQ1 <- as.numeric(dataAugmented$issFriendQ1)
dataAugmented$issFriendQ2 <- as.numeric(dataAugmented$issFriendQ2)
dataAugmented$issFriendQ3 <- as.numeric(dataAugmented$issFriendQ3)
dataAugmented$issFriendQ4 <- as.numeric(dataAugmented$issFriendQ4)
dataAugmented$issFriendQ5 <- as.numeric(dataAugmented$issFriendQ5)
dataAugmented$aceQ1 <- as.numeric(dataAugmented$aceQ1)
dataAugmented$aceQ2 <- as.numeric(dataAugmented$aceQ2)
dataAugmented$aceQ3 <- as.numeric(dataAugmented$aceQ3)
dataAugmented$aceQ4 <- as.numeric(dataAugmented$aceQ4)
dataAugmented$aceQ5 <- as.numeric(dataAugmented$aceQ5)
dataAugmented$aceQ6 <- as.numeric(dataAugmented$aceQ6)
dataAugmented$aceQ7 <- as.numeric(dataAugmented$aceQ7)
dataAugmented$aceQ8 <- as.numeric(dataAugmented$aceQ8)
dataAugmented$aceQ9 <- as.numeric(dataAugmented$aceQ9)
dataAugmented$aceQ10 <- as.numeric(dataAugmented$aceQ10)
dataAugmented$gritQ1 <- as.numeric(dataAugmented$gritQ1)
dataAugmented$gritQ2 <- as.numeric(dataAugmented$gritQ2)
dataAugmented$gritQ3 <- as.numeric(dataAugmented$gritQ3)
dataAugmented$gritQ4 <- as.numeric(dataAugmented$gritQ4)
dataAugmented$gritQ5 <- as.numeric(dataAugmented$gritQ5)
dataAugmented$gritQ6 <- as.numeric(dataAugmented$gritQ6)
dataAugmented$gritQ7 <- as.numeric(dataAugmented$gritQ7)
dataAugmented$gritQ8 <- as.numeric(dataAugmented$gritQ8)
```

## Adjust Value Assignments
The following code chunk adjusts the values assigned to fixed pre-determined response choices for select variables and recodes blanks with `NA` in the dataset.
```{r adjust-values, message = FALSE, warning = FALSE}
dataAugmented$ageNow <- (as.numeric(dataAugmented$ageNow)+11)
dataAugmented$ageRelease <- (as.numeric(dataAugmented$ageRelease)+11)

# Recode variables with `yes` and `no` response options as 1 and 0 respectively
dataAugmented %>%
  mutate(savings = ifelse (savings == 1,1,0)) %>%
  mutate(helpDuring = ifelse(helpDuring == 1,1,0)) %>%
  mutate(helpAfter = ifelse(helpAfter == 1,1,0)) -> dataAugmented
#  mutate(savings = case_when(
#    as.numeric(savings) == 1 ~ 1,
#    as.numeric(savings) == 2 ~ 0,
#    TRUE ~ savings)
#  )%>%
#  mutate(helpDuring = case_when(
#    as.numeric(helpDuring) == 1 ~ 1,
#    as.numeric(helpDuring) == 2 ~ 0,
#    TRUE ~ helpDuring)
#  )%>%
#  mutate(helpAfter = case_when(
#    as.numeric(helpDuring) == 1 ~ 1,
#    as.numeric(helpDuring) == 2 ~ 0,
#    TRUE ~ helpAfter)
#    )-> dataAugmented

# Recode instances of missing data with `NA`
dataAugmented[dataAugmented==""] <- NA
```

## Create Variables
The following code chunk creates additional variables needed for the study.
```{r scales, message = FALSE, warning = FALSE}
dataAugmented %>%
  # Create variable to consolidate `ethnicityRace` classifications
  # Coding: 1 - black, 2 - white, 3 - other
  mutate(ethnicityRaceGrouped = case_when(
    ethnicityRace == 4 ~ 1,
    ethnicityRace == 9 ~ 2,
    TRUE ~ 3)
  ) %>%
  
  # Create variable to consolidate `religiousAffiliation`
  # Coding: 1 - religious or spiritual, 2 - not religious or spiritual, 3 - prefer not to say
  mutate(religiousAffilGrouped = case_when(
    between (religiousAffiliation, 3, 11) | religiousAffiliation==12 ~ 1,
    between (religiousAffiliation, 1, 2) ~ 2,
    religiousAffiliation == 13 ~ 13)
    ) %>%
  
  # Create variable to consolidate `typeHometown`
  # Coding: 1 - less than or equal to 50,000; 2 - greater than 50,000
  mutate(typeHometownGrouped = case_when(
    between (typeHometown, 1, 2) ~ 1,
    typeHometown == 3 ~ 2)
    ) %>%
  
  # Calculate score for multidimensional scale for percieved social support
  mutate (mspssScore = case_when(
    !is.na(pssQ1) & !is.na(pssQ2) & !is.na(pssQ3) & !is.na(pssQ4) & !is.na(pssQ5) & !is.na(pssQ6) & !is.na(pssQ7) &
      !is.na(pssQ8) & !is.na(pssQ9) & !is.na(pssQ10) & !is.na(pssQ11) & !is.na(pssQ12) ~
      (pssQ1+pssQ2+pssQ3+pssQ4+pssQ5+pssQ6+pssQ7+pssQ8+pssQ9+pssQ10+pssQ11+pssQ12)/12)
    )%>%
  
  # Calculate score for family instrumental social support
  mutate (issFamilyScore = case_when(
    !is.na(issFamilyQ1) & !is.na(issFamilyQ2) & !is.na(issFamilyQ3) & !is.na(issFamilyQ4) & !is.na(issFamilyQ5)
    ~ (issFamilyQ1+issFamilyQ2+issFamilyQ3+issFamilyQ4+issFamilyQ5)/5)
  )%>%
  
  # Calculate score for friend instrumental social support
  mutate (issFriendScore = case_when(
    !is.na(issFriendQ1) & !is.na(issFriendQ2) & !is.na(issFriendQ3) & !is.na(issFriendQ4) & !is.na(issFriendQ5)
    ~ (issFriendQ1+issFriendQ2+issFriendQ3+issFriendQ4+issFriendQ5)/5)
  )%>%
  
  # Calculate score for total instrumental social support
  mutate(issScore = case_when(
    !is.na(issFriendScore) & !is.na(issFamilyScore) ~ (issFamilyScore+issFriendScore)/2)
  )%>%
  
  # Calculate score for adverse childhood experience
  mutate (aceScore = case_when (
    !is.na(aceQ1) & !is.na(aceQ2) & !is.na(aceQ3) & !is.na(aceQ4) & !is.na(aceQ5) & !is.na(aceQ6) & !is.na(aceQ7) &
      !is.na(aceQ8) & !is.na(aceQ9) & !is.na(aceQ10) ~ 
      (aceQ1+aceQ1+aceQ2+aceQ3+aceQ4+aceQ5+aceQ6+aceQ7+aceQ8+aceQ9+aceQ10))
    ) %>%
  
  # Calculate socre for grit
  mutate (gritQ1R = 6-gritQ1, gritQ3R = 6-gritQ3, gritQ5R = 6-gritQ5, gritQ6R = 6-gritQ6) %>%
  mutate (gritScore = case_when(
    !is.na(gritQ1R) & !is.na(gritQ2) & !is.na(gritQ3R) & !is.na(gritQ4) & !is.na(gritQ5R) &
      !is.na(gritQ6R) & !is.na(gritQ7) & !is.na(gritQ8) ~ 
      ((gritQ1R+gritQ2+gritQ3R+gritQ4+gritQ5R+gritQ6R+gritQ7+gritQ8)/8)
    ))%>%
  
  # Calculate total incarceration time in months
  mutate (incarcerationTime = ((incarcerationYears-1)*12)+(incarcerationMonths-1)) %>%
  
  # Calculate total time released
  mutate (timeReleased = ageNow - ageRelease) %>%
  
  # Impute 1 (coding for 0%) for `NA` in `selfEmployment` variable
  mutate (selfEmployment = case_when(
    is.na(selfEmployment) ~ 1)
    )%>%
  
  # Calculate interval variable for poverty ratio
  mutate (povertyRatio = if(as.numeric(incomeLastYr) <= 6) {
    (12490+((as.numeric(incomeLastYr)-1))*4420)/(8070+(as.numeric(householdSize)*4420))
    } else {
      if(as.numeric(incomeLastYr) == 7) {
        (59590/(8070+(as.numeric(householdSize)*4420)))
      } else {
        if(as.numeric(incomeLastYr) == 8) {
          (62500/(8070+(as.numeric(householdSize)*4420)))
           } else {
             if(as.numeric(incomeLastYr) == 9) {
               (87500/(8070+(as.numeric(householdSize)*4420)))
             } else {
               if(as.numeric(incomeLastYr) == 10) {
                 (125000/(8070+(as.numeric(householdSize*4420))))
               } else {
                 (150000/(8070+(as.numeric(householdSize)*4420)))
               }
             }
          }
      }
    }
    ) %>%
  
  # Calculate change in income
  mutate(incomeChange = incomeLastYr - incomeInitial) %>%
  
  # Determine highest education level
  mutate(educationHighest = sapply(strsplit(as.character(educationLevels), split = ","), 
         function(x) max(as.numeric(x)))) %>%
  
  # Calculate binary variable indicating food insecurity
  mutate(foodInsecurity = case_when(
    foodSecurityQ1 == 2|foodSecurityQ1 == 3|foodSecurityQ2 == 2|foodSecurityQ2 == 3 ~ 1,
    foodSecurityQ1 == 1 & foodSecurityQ2 == 1 ~ 0)
    )%>%
  
  # Calculate binary variable indicating use of any social services
  mutate (programsUsedBinary = case_when(
    is.na(programsUsed) ~ 0,
    !is.na(programsUsed) ~ 1)
    ) %>%
  
  # Determine number of public and social services used after release
  mutate(countServicesAfter = count.fields(textConnection(as.character(programsUsed)), sep = ",")) %>%
  
  # Determine number of social services used during incarceration
  mutate(countHelpDuring = count.fields(textConnection(as.character(helpDuringType)), sep = ",")) %>%
  
  # Determine number of social services used after release
  mutate(countHelpAfter = count.fields(textConnection(as.character(helpAfterType)), sep = ",")) -> dataAugmented
```

## Select Cases
The following code chunk removes cases where the respondent was part of a protected population or did not provide informed consent. The original data for the variables was not recoded. It then removes duplicate cases.
```{r select-cases, message = FALSE, warning = FALSE}
# Case 2 (responseID R_3ea7BOzjDjNiJYr) and Case 15 (responseID R_1ouyBTWoxJv4FkS) are the same.  Remove case 2 in the final analysis.
dataAugmented %>%
  subset(currentlyIncarcerated=2) %>%
  subset(currentlyDetained=2) %>%
  subset(residentialTreatment=2) %>%
  subset(involCommitment=2) %>%
  dplyr::distinct(.,.keep_all = TRUE) -> dataClean
```

## Evaluate Missing Data
The following code chunk evaluates missing data.
```{r missing-data, message = FALSE, warning = FALSE}
miss_var_summary(dataClean, order = TRUE)
miss_case_summary(dataClean, order = TRUE)
```

## Calculate Descriptive Statistics
The following code calculates descriptive statistics for select variables of interest.
```{r descriptive-stats, message = FALSE, warning = FALSE}
print("Number of cases") 
nrow(dataClean)

print("gender; 1 - male, 2 - female, 3 - other, 4 - undisclosed")
print("values")
table(as.numeric(dataClean$gender))
print("proportions")
prop.table(table(as.numeric(dataClean$gender)), margin = NULL)

print("ethnicityRace; 1 - East Asian, 2 - Central Asian, 3 - Southern Asian, 4 - Black or African-American,
      5 - Hispanic or Latino, 6 - Middle Eastern or North African, 7 - Native American or Alaska,
      8 - Native Hawaiian or Pacific Islander, 9 - White or European, 10 - Other")
print("values")
table(as.numeric(dataClean$ethnicityRace))
print("proportions")
prop.table(table(as.numeric(dataClean$ethnicityRace)), margin = NULL)

print("religiousAffiliation; 1 - Atheism, 2 - Agnoticism, 3 - Buddhist, 4 - Eastern Orthodox, 5 - Hindu,
      6 - Islam, 7 - Jewish, 8 - Latter Day Saints, 9 - Non-denominational, 10 - Protestant,
      11 - Roman Catholic, 12 - Other, 13 - prefer not to answer")
print("values")
table(as.numeric(dataClean$religousAffiliation))
print("proportions")
prop.table(table(as.numeric(dataClean$religiousAffiliation)), margin = NULL)

print("educationLevels; 1 - high schoor or GED, 2 - trade school, 3 - some college, 4 - associate,
      5 - bachelor, 6 - master, 7 - doctorate")
print("values")
table(as.numeric(dataClean$educationLevels))
print("proportions")
prop.table(table(as.numeric(dataClean$educationLevels)), margin = NULL)

print("whenHighestEd; 1 - before, 2 - during, 3 - after")
print("values")
table(as.numeric(dataClean$whenHighestEd))
print("proportions")
prop.table(table(as.numeric(dataClean$whenHighestEd)), margin = NULL)

print("typeHometown; 1 - rural, 2 - suburban, 3 - urban")
print("values")
table(as.numeric(dataClean$typeHometown))
print("proportions")
prop.table(table(as.numeric(dataClean$typeHometown)), margin = NULL)

print("ageNow; ratio data (years)")
describe(as.numeric(dataClean$ageNow), na.rm = TRUE)

print("ageRelease; ratio data (years)")
describe(as.numeric(dataClean$ageRelease), na.rm = TRUE)

print("incarcerationTime; ratio data (months)")
summary(as.numeric(dataClean$incarcerationTime))

print("incomeInitial; ordinal data")
print("values")
table(as.numeric(dataClean$incomeInitial))
print("proportions")
prop.table(table(as.numeric(dataClean$incomeInitial)), margin = NULL)

print("incomeLastYr; ordinal data")
print("values")
table(as.numeric(dataClean$incomeLastYr))
print("proportions")
prop.table(table(as.numeric(dataClean$incomeLastYr)), margin = NULL)

print("selfEmployment; 1 - 0%, 2 - 1-25%, 3 - 26-50%, 4 - 51-75%, 5 - 75-99%, 6 - 100%")
print("values")
table(as.numeric(dataClean$selfEmployment))
print("proportions")
prop.table(table(as.numeric(dataClean$selfEmployment)), margin = NULL)

print("foodInsecurity; 1 - yes, 0 - no")
print("values")
table(as.numeric(dataClean$foodInsecurity))
print("proportions")
prop.table(table(as.numeric(dataClean$foodInsecurity)), margin = NULL)
```

## Compute Scale Reliabilities
The following code chunk evaluates the reliability for the MSPSS, ACE, and grit scales by calculating the Cronbach's alpha for each variable.
```{r reliabilities, message = FALSE, warning = FALSE, vspace="1em"}
reliabilityMSPSS <- cronbach (subset(dataClean, select=c(pssQ1, pssQ2, pssQ3, pssQ4, pssQ5, pssQ6, pssQ7, pssQ8, pssQ9, pssQ10, pssQ11, pssQ12)))
print("Reliability for MSPSS")
reliabilityMSPSS

reliabilityISS <- cronbach (subset(dataClean, select=c(issFamilyQ1, issFamilyQ2, issFamilyQ3, issFamilyQ4, issFamilyQ5, 
                                                            issFriendQ1, issFriendQ2, issFriendQ3, issFriendQ4, issFriendQ5)))
print("Reliability for Total Instrumental Social Support")
reliabilityISS

reliabilityISSfamily <- cronbach (subset(dataClean, select=c(issFamilyQ1, issFamilyQ2, issFamilyQ3, issFamilyQ4, issFamilyQ5)))
print("Reliability for Family Instrumental Social Support")
reliabilityISSfamily

reliabilityISSfriend <- cronbach (subset(dataClean, select=c(issFriendQ1, issFriendQ2, issFriendQ3, issFriendQ4, issFriendQ5)))
print("Reliability for Family Instrumental Social Support")
reliabilityISSfriend

reliabilityACE <- cronbach (subset(dataClean, select=c(aceQ1, aceQ2, aceQ3, aceQ4, aceQ5, aceQ6, aceQ7, aceQ8, aceQ9, aceQ10)))
print("Reliability for ACE Scale")
reliabilityACE

reliabilityGrit <- cronbach (subset(dataClean, select=c(gritQ1R, gritQ2, gritQ3R, gritQ4, gritQ5R, gritQ6R, gritQ7, gritQ8)))
print("Reliability for Short Grit Scale")
reliabilityGrit

reliabilityHungerVitalSigns <- cronbach (subset(dataClean, select=c(foodSecurityQ1,foodSecurityQ2)))
print("Reliability for Hunger Vital Signs")
reliabilityHungerVitalSigns
```

## Perform Comparison of Means of Grit
The following code chunk performs calculations for comparison of means of the primary variables of interest.
```{r means, message = FALSE, warning = FALSE}
# Comparison of means for `gritScore` grouped by various operationalizations of success using ANOVA
# Coding for `gritscore`: values range from 1 (low grit) to 8 (high grit)
# Null hypothesis: the means of the different groups are the same
# Alternative hypothesis: the sample mean of at least one group is not equal to the others


# Comparison of `gritScore` for cases grouped by `incomeLastYr` using ANOVA
# Coding for `incomeLastYr` (Qualitrics Q22): ordinal ranging 1 - 10
groupIncome <- group_by(dataClean, incomeLastYr)

groupIncome %>%
  summarise(
    count = n(),
    mean = mean(gritScore, na.rm = TRUE),
    sd = sd(gritScore, na.rm = TRUE)
  )

aovGroupIncome <- aov(gritScore ~ as.numeric(incomeLastYr), data = groupIncome)
summary(aovGroupIncome)

# Comparison of `gritScore` for cases grouped by `povertyLevel` using ANOVA
# Coding for `povertyLevel` (created variable): 1 - above, 0 - at or below
dataClean %>%
  mutate(povertyLevel = if (povertyRatio >1) {1} else {0})-> dataClean

groupPoverty <- group_by(dataClean, povertyLevel)

groupPoverty %>%
  summarise(
    count = n(),
    mean = mean(gritScore, na.rm = TRUE),
    sd = sd(gritScore, na.rm = TRUE)
  )

aovGroupPoverty <- aov(gritScore ~ as.numeric(povertyLevel), data = groupPoverty)
summary(aovGroupPoverty)

# Comparison of `gritScore` for cases grouped by `managingFinancially` using ANOVA
# Coding for `managingFinancially` (Qualitrics Q23): 1 - not able, 2 - just able, 3 - comfortable no savings, 4 - comfortable and saving
groupFinStat <- group_by(dataClean, managingFinancially)

groupFinStat %>%
  summarise(
    count = n(),
    mean = mean(gritScore, na.rm = TRUE),
    sd = sd(gritScore, na.rm = TRUE)
  )

aovGroupFinStat <- aov(gritScore ~ as.numeric(managingFinancially), data = groupFinStat)
summary(aovGroupFinStat)

# Comparison of `gritScore` for cases grouped by `savings` using ANOVA
# Coding for `savings` (Qualitrics Q24): 1 - yes, 2 - no
groupSavings <- group_by(dataClean, savings)

groupSavings %>%
  summarise(
    count = n(),
    mean = mean(gritScore, na.rm = TRUE),
    sd = sd(gritScore, na.rm = TRUE)
  )

aovGroupSavings <- aov(gritScore ~ as.numeric(savings), data = groupSavings)
summary(aovGroupSavings)
```
## Comparison of Means for Categorical Variables
The following code chunk performs comparison of means for select categorical variables of interest.
```{r chi-square, message = FALSE, warning = FALSE}
# Null hypothesis: The two variables are independent
# Alternative hypothesis: The two variables relate to each other

print("`typeHometown` and `foodInsecurity`")
ChiSqr01 <-  chisq.test(dataClean$typeHometown, dataClean$foodInsecurity)
ChiSqr01

print("`whenHighestEd` and `incomeLastYr`")
ChiSqr02 <- chisq.test(dataClean$whenHighestEd, dataClean$incomeLastYr)
ChiSqr02
```

## Perform Correlational Analysis
The following code chunk performs correlational analysis for the primary variables of interest.
```{r correlation, message = FALSE, warning = FALSE}
dataClean %>%
  dplyr::select(incomeLastYr, incomeChange, povertyRatio, managingFinancially, savings, programsUsedBinary,
                countServicesAfter, countHelpDuring, countHelpAfter, educationHighest, ageNow, ageRelease,
                incarcerationTime, timeReleased, mspssScore, issScore, aceScore, gritScore 
                ) -> correlationDataset 
col_names <- names(correlationDataset)
correlationDataset[,col_names] <- lapply(correlationDataset[,col_names], as.numeric)
correlation <- cor(correlationDataset, use="pairwise.complete.obs")
correlation
```

## Perform Regression Analysis
The following code chunk performs a multiple regression using select variables of interest.
```{r regression, message = FALSE, warning = FALSE}
# Multiple regression using `incomeLastYr` as dependent variable
regressionIncome <- lm(as.numeric(incomeLastYr) ~ as.numeric(programsUsedBinary) + as.numeric(educationHighest) +
                         as.numeric(ageRelease) + as.numeric(timeReleased) + mspssScore + issScore + aceScore + gritScore + 
                         gender + as.factor(ethnicityRace) + as.factor(typeHometown), 
                        data = dataClean, na.action = na.omit)
summary(regressionIncome)

# Multiple regression using `incomeRelease` as dependent variable
regressionIncomeInitial <- lm(as.numeric(incomeInitial) ~ as.numeric(programsUsedBinary) + as.numeric(educationHighest) +
                         as.numeric(ageRelease) + as.numeric(timeReleased) + mspssScore + issScore + aceScore + gritScore + 
                         gender + as.factor(ethnicityRace) + as.factor(typeHometown), 
                        data = dataClean, na.action = na.omit)
summary(regressionIncomeInitial)

# Multiple regressiong using `povertyRatio` as dependent variable
regressionPoverty <- lm(as.numeric(povertyRatio) ~ as.numeric(programsUsedBinary) + as.numeric(educationHighest) +
                          as.numeric(ageRelease) + as.numeric(timeReleased) + mspssScore + issScore + aceScore + gritScore + 
                          gender + as.factor(ethnicityRace) + as.factor(typeHometown),  
                        data = dataClean, na.action = na.omit)
summary(regressionPoverty)


# Multiple regression using `managingFinancially` as dependent variable
regressionFinStat <- lm(as.numeric(managingFinancially) ~ as.numeric(programsUsedBinary) + as.numeric(educationHighest) +
                          as.numeric(ageRelease) + as.numeric(timeReleased) + mspssScore + issScore + aceScore + gritScore + 
                          gender + as.factor(ethnicityRace) + as.factor(typeHometown),  
                        data = dataClean, na.action = na.omit)
summary(regressionFinStat)


# Binary logistic regression using `savings` as dependent variable
# Select variables
dataClean %>%
  dplyr::select(incomeLastYr, povertyRatio, povertyLevel, managingFinancially, savings, programsUsedBinary, educationHighest, 
                ageRelease, timeReleased, gender, mspssScore, issScore, aceScore, gritScore) -> dataBinary

# Remove cases with missing data
dataBinary %>%
  drop_na() -> dataBinary

logitSavings <- glm(savings ~ programsUsedBinary + managingFinancially + educationHighest + ageRelease + timeReleased + 
                      mspssScore + issScore + aceScore + gritScore + gender, 
                     data = dataBinary, family = binomial, na.action = na.omit)
summary(logitSavings)

# Raise e to the coefficients
print(exp(coef(logitSavings)))

# Obtain various pseudo R-squared measures
print(pR2(logitSavings))

# Confidence intervals for the coefficients
print(exp(confint(logitSavings, level = 0.95)))

# Hosmer-Lemeshow Goodness of Fit Test
# Null hypothesis: the model is a good fit for the data
# Alternative hypothesis: the model is NOT a good fit for the data
HosLemLogitSavings <- hoslem.test(dataBinary$savings, 
                              fitted(logitSavings), g=10)
print(HosLemLogitSavings)
print(cbind(HosLemLogitSavings$expected, HosLemLogitSavings$observed))
```

## Save Data
The following code chunk saves the cleaned data used for the analysis. 
```{r save-analysis, message = FALSE, warning = FALSE}
write.csv(dataClean, here("Data","DataClean","SOC5800_Data_NumericValues_Clean_CSV.csv"), append = FALSE)
```